Appendix B: Code for analysis comparing to historical data

### 2018-01-29 appendix/appendixB.Rmd
appendixB tidy

Computes model parameter estimates for selected stocks in RAM using NIMBLE. estimate $B_{MSY}$

```{r knitr, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE, dev="cairo_pdf")
```

@@

```{r plotting, include = FALSE}
## This chunk is optional, just tweaks plot appearance / fonts
library(ggthemes)
library(Cairo)
library(extrafont)
library(hrbrthemes)

@@

```{r}
if(!file.exists("ramlegacy.zip")){
download.file("https://depts.washington.edu/ramlegac/wordpress/databaseVersions/RLSADB_v3.0_(assessment_data_only)_excel.zip",
              "ramlegacy.zip")
}
```

@@

```{r}
## Model does not estimate sigma_m; data is insufficient to do so.
## Note that RAM stock estimates are not always direct measurements, can be artifically smooth, underestimating sigma_g

gs_code  <- nimble::nimbleCode({

  r ~ dunif(0, 2)
  K ~ dunif(0, 2)
  sigma ~ dunif(0, 1)

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K) - min(a[t],x[t])
    x[t+1] ~ dnorm(mu[t], sd = sigma)
  }

})

@@

  samples <- as.data.frame(as.matrix(Cmcmc$mvSamples))
  burnin <- 1:(0.05 * dim(samples)[1]) # drop first 5%
  samples <- samples[-burnin,1:(length(inits) - 1)] # drop raised vars, burnin

  #gather(samples) %>% ggplot() + geom_density(aes(value)) + facet_wrap(~key, scale='free')

@@

<!--
Do the estimates make sense?  Consider simulation from models, under this historical harvests, compared to historical trajectories

```{r}
examples %>% ungroup() %>%
  select(year, scaled_biomass, scaled_catch, commonname) %>%
  gather(stock, biomass, -year, -commonname) %>%
  ggplot(aes(year, biomass, col=stock)) + 
  geom_line() + facet_wrap(~commonname, scales = "free")
```

-->

@@

# Decision Policies

@@

## All parameter values combinations for which we want solutions

@@

## Create the models

@@

@@

```{r message=FALSE}
meta <- meta_from_log(data.frame(model="gs"), log_dir)  %>% 
  mutate(scenario = as.character(scenario)) %>%
  left_join(
    ## use the estimated sigma_m, r, K pars recorded in the logged meta.csv.  Technically
    ## these should be the same as the above, but without random seed re-running the nimble
    ## code but not re-running these can create mis-matches.  
    select(meta, sigma_m, commonname, scenario),   
    by = c("scenario", "sigma_m", "commonname")) %>% 
  arrange(scenario)
alphas <- alphas_from_log(meta, log_dir)
#models <- models_from_log(meta)
```

@@

  tibble(S_star, F_MSY = MSY / B_MSY, F_PGY = 0.8 * F_MSY, 
         commonname = P$commonname, N = P$N)
}

policy_pars <- 
  pars %>% 
  rowwise() %>% 
  do(statics(.))
```  
  Convert example data into discrete index space.

```{r}
index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
## repeats each series for each static model
ex <- examples %>%
  mutate(biomass = index(scaled_biomass, states),          
         catch = index(scaled_catch, actions))  %>% 
  left_join(policy_pars) %>% left_join(pars) %>%
  ungroup() 
ex
```

```{r} 
det_f <- function(S_star, r, K, i) index(pmax(gs(r[[1]],K[[1]])(states,0) - S_star[[1]],0), actions)[i]
msy_f <- function(F_MSY, i) index(states * F_MSY[[1]], actions)[i]
pgy_f <- function(F_PGY, i) index(states * F_PGY[[1]], actions)[i]

rescale <- function(x, N) states[x]*N

historical <- ex %>% 
  group_by(commonname) %>% 
  mutate(det =  det_f(S_star, r, K, biomass),
         msy =  msy_f(F_MSY, biomass),
         pgy =  pgy_f(F_PGY, biomass)) %>%
  select(year, biomass, catch, det, msy, pgy, commonname, N) %>%
  gather(model, stock, -year, -commonname, -N) %>%
  mutate(stock = states[stock] * N) %>% 
  select(-N)
```

## Plots

```{r}
historical %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = c("black", "grey", "#D9661F", "#00B0DA", "#CFDD45")) +
  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
```

## POMDP historical

@@

@@

```{r}
sims <- pomdp_sims %>% 
  mutate(optimal = states[optimal] * N) %>% # original scale
  select(year, optimal, commonname, sigma_m) %>%
  rename(stock = optimal) %>% 
  ## treat each sigma_m value as separate 'model'
  mutate(sigma_m = as.factor(sigma_m)) %>%
  mutate(model = recode(sigma_m, 
                        "0" = "Reed",
                        "0.1" = "POMDP", 
                        "0.15" = "POMDP_0.15")) %>%
  select(-sigma_m) %>%
  bind_rows(historical)

@@

```{r}
ram_ex <- read_csv(file.path(log_dir, "ram_ex.csv"))

ram_ex %>%
  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy", "msy")) %>%
   mutate(model = fct_recode(model,
                             "MSY" = "msy",
                             "CE" = "Reed",
                             "TAC" = "pgy"
                             )) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = colors) +
  facet_wrap(~commonname, scales = "free", ncol=1)  
```

>>>>

Computes model parameter estimates for selected stocks in RAM using NIMBLE.

```{r knitr, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=TRUE, dev="cairo_pdf")
```

@@

```{r plotting, include = FALSE}
## This chunk is optional, just tweaks plot appearance / fonts
library(printr)
library(ggthemes)
library(Cairo)
library(extrafont)
library(hrbrthemes)

@@

```{r}
if(!file.exists("ramlegacy.zip")){
download.file(paste0(
  "https://depts.washington.edu/ramlegac/wordpress/databaseVersions/",
  "RLSADB_v3.0_(assessment_data_only)_excel.zip"),
  "ramlegacy.zip")
}
```

@@

```{r}
## Model does not estimate sigma_m; data is insufficient to do so.
gs_code  <- nimble::nimbleCode({
  r ~ dunif(0, 2)
  K ~ dunif(0, 2)
  sigma ~ dunif(0, 1)

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K) - min(a[t],x[t])
    x[t+1] ~ dnorm(mu[t], sd = sigma)
  }
})

@@

  samples <- as.data.frame(as.matrix(Cmcmc$mvSamples))
  burnin <- 1:(0.05 * dim(samples)[1]) # drop first 5%
  samples <- samples[-burnin,1:(length(inits) - 1)] # drop raised vars, burnin
  #gather(samples) %>% ggplot() + geom_density(aes(value)) + facet_wrap(~key, scale='free')

@@

```{r include = FALSE, eval = FALSE}
## Do the estimates make sense?  Consider simulation from models, 
## under this historical harvests, compared to historical trajectories
examples %>% ungroup() %>%
  select(year, scaled_biomass, scaled_catch, commonname) %>%
  gather(stock, biomass, -year, -commonname) %>%
  ggplot(aes(year, biomass, col=stock)) + 
  geom_line() + facet_wrap(~commonname, scales = "free")
```

@@

# Calculations of the Decision Policies for Historical Data

@@

Consider all parameter values combinations for which we want solutions (both species at each of three possible levels of measurement uncertainty; though we will focus on the `0.1` level for simplicity as overall pattern is the same at `0.15`):

@@

Create the model matrices (transition, observation, and reward matrix):

@@

Here's the slowest part: computing POMDP alpha vectors.

@@

```{r message=FALSE}
meta <- meta_from_log(data.frame(model="gs"), log_dir)  %>% 
  mutate(scenario = as.character(scenario)) %>%
  left_join(
    select(meta, sigma_m, commonname, scenario),   
    by = c("scenario", "sigma_m", "commonname")) %>% 
  arrange(scenario)
alphas <- alphas_from_log(meta, log_dir)
#models <- models_from_log(meta)
```

@@

  tibble(S_star, F_MSY = MSY / B_MSY, F_TAC = 0.8 * F_MSY, 
         commonname = P$commonname, N = P$N)
}

policy_pars <- 
  pars %>% 
  rowwise() %>% 
  do(statics(.))
```

Convert example data into discrete index space.

```{r}
index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
## repeats each series for each static model
ex <- examples %>%
  mutate(biomass = index(scaled_biomass, states),          
         catch = index(scaled_catch, actions))  %>% 
  left_join(policy_pars) %>% 
  left_join(pars) %>%
  ungroup() 
```

Static policy calculations:

```{r} 
CE_f <- function(S_star, r, K, i) 
  index(pmax(gs(r[[1]],K[[1]])(states,0) - S_star[[1]],0), actions)[i]
MSY_f <- function(F_MSY, i) index(states * F_MSY[[1]], actions)[i]
TAC_f <- function(F_TAC, i) index(states * F_TAC[[1]], actions)[i]
rescale <- function(x, N) states[x]*N

historical <- ex %>% 
  group_by(commonname) %>% 
  mutate(CE =  CE_f(S_star, r, K, biomass),
         MSY =  MSY_f(F_MSY, biomass),
         TAC =  TAC_f(F_TAC, biomass)) %>%
  select(year, biomass, catch, CE, MSY, TAC, commonname, N) %>%
  gather(model, stock, -year, -commonname, -N) %>%
  mutate(stock = states[stock] * N) %>% 
  select(-N)
```

Compute POMDP policy for historical data:

@@

Join records:

@@

```{r}
sims <- pomdp_sims %>% 
  mutate(optimal = states[optimal] * N) %>% # original scale
  select(year, optimal, commonname, sigma_m) %>%
  rename(stock = optimal) %>% 
  ## treat each sigma_m value as separate 'model'
  mutate(sigma_m = as.factor(sigma_m)) %>%
  mutate(model = recode(sigma_m, 
                        "0" = "CE",
                        "0.1" = "POMDP", 
                        "0.15" = "POMDP_0.15")) %>%
  select(-sigma_m) %>%
  bind_rows(historical)

@@

Final plot, as in paper but including MSY:

```{r}
ram_ex <- read_csv(file.path(log_dir, "ram_ex.csv"))
ram_ex %>%
  filter(model %in% c("biomass", "catch", "POMDP", "CE", "TAC", "MSY")) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = colors) +
  facet_wrap(~commonname, scales = "free", ncol=1)  
```
